---
title: "FitBit"
author: "Abir"
date: "24/02/2022"
output: html_document
---
## ASK PHASE ##

**Buisness Task**

* Analyze Fitbit user data to understand how consumers are using their tracker.
* Provide recommendations for the Bellabeat marketing team.

## PREPARE ##
**About the data**

This dataset was downloaded from Kaggle.

**Credibility of data**

Using the ROCCC method to determine the credibility of the data.ROCCC is reliability, 
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(skimr)
library(janitor)
```
**Loading and naming the data that we are going to work**
```{r}
setwd("C:/Users/HP/Desktop/Notes/DATA ANALYST/Project/Bellabeat/archive/Fitabase Data")
daily_activity <- read.csv("dailyActivity_merged.csv")
weight_log <- read.csv("weightLogInfo_merged.csv")
sleep_log <- read.csv("sleepDay_merged.csv")
```
## PROCESS ##
```{r}
colnames(daily_activity)
```
```{r}
colnames(weight_log)
```
```{r}
colnames(sleep_log)
```
```{r}
head(daily_activity)
```
```{r}
head(weight_log)
```
```{r}
head(sleep_log)
```
Counting of how many distinct values are in the datasets' ID columns
```{r}
n_distinct(daily_activity$Id)
```
```{r}
n_distinct(weight_log$Id)
```
```{r}
n_distinct(sleep_log$Id)
```
From the number of distinct ID's enteres in three datasets, there were much people who manually entered their data.
Now, checking the duplicates.
```{r}
nrow(daily_activity)
```
```{r}
nrow(weight_log)
```
```{r}
nrow(sleep_log)
```
Now,
```{r}
nrow(daily_activity[duplicated(daily_activity),])
```
```{r}
nrow(weight_log[duplicated(weight_log),])
```
```{r}
nrow(sleep_log[duplicated(sleep_log),])
```
There are 3 duplicates in sleep_log dataframe. So we create new sleep_log_new data frame
without any duplicates
```{r}
sleep_log_new <- unique(sleep_log)
```
Now, checking sleep_log_new for duplicates
```{r}
nrow(sleep_log_new[duplicated(sleep_log_new),])
```
```{r}
nrow(sleep_log_new)
```
## Analyze ##
Now that the data is loaded and clean, let's look at the data's summary statistics
```{r}
daily_activity %>%
  select(TotalSteps,
         TotalDistance,
         FairlyActiveMinutes,
         VeryActiveDistance,
         SedentaryMinutes,
         Calories) %>%
  summary()
```
```{r}
sleep_log_new %>%
  select(TotalSleepRecords,
         TotalMinutesAsleep,
         TotalTimeInBed)%>%
  summary()
```
```{r}
weight_log %>%
  select(WeightPounds,
         BMI)%>%
  summary()
```
**Observations of analysis**

daily_activity:

* The average total daily steps taken by users were 7,360.
* On average, users burned 2,304 calories daily.
* Average very active Minutes were 21.16 minutes.
* Average Sedentary Minutes were 991.2 minutes.

sleep_log_new:

* Users slept on average of 419.2 minnutes, or around 7 hors.
* Users time asleep and time in bed was fairly similar.

weight_log:

* The data in weight_log in very less.
* Users are not entering their data into the app.

## Visualizations ##
**Relationships between Total Steps and Sedentary Minutes**
```{r}
ggplot(data=daily_activity)+
  geom_point(mapping = aes(x=TotalSteps, y=SedentaryMinutes, color = Calories))+
  geom_smooth(mapping = aes(x=TotalSteps, y=SedentaryMinutes))+
  labs(title="Relationship Between Total Steps and Sedentary Minutes", x="Total Steps",y="Sedentary Minutes")
```
**Relationship between total minutes asleep and total times in bed**
```{r}
ggplot(data=sleep_log_new)+
  geom_point(mapping=aes(x=TotalMinutesAsleep, y=TotalTimeInBed,))+
  geom_smooth(mapping=aes(x=TotalMinutesAsleep, y=TotalTimeInBed,))+
  labs(title = "Relationship between total minutes asleep and total times in bed", x = "TotalMinutesAsleep", y = "TotalTimeInBed") 
```

```{r}
ggplot(data = daily_activity) +
  geom_point(mapping=aes(x=TotalSteps, y=Calories, color= VeryActiveMinutes)) +
  labs(title = "Relationship Between Total Steps and Calories", x = "Total Steps", y = "Calories Burned") 
```
**Relationship between Sedentary Minutes and Calories**
```{r}
ggplot(data = daily_activity)+
  geom_point(mapping = aes(x=SedentaryMinutes,y=Calories, color = "violet"))+
  geom_smooth(mapping = aes(x=SedentaryMinutes,y=Calories))+
  labs(title = "Relationship between Sedentary Minutes and Calories",x="Sedentary Minutes",y="Calories Burned")
```
**Relationship between Very active minutes and calories**
```{r}
ggplot(data = daily_activity) +
  geom_point(mapping = aes(x= VeryActiveMinutes, y = Calories, color = "red")) + 
  geom_smooth(mapping=aes(x=VeryActiveMinutes, y=Calories)) +
  labs(title = "Relationship Between Very Active Minutes and Calories", x = "Very Active Minutes", y = "Calories Burned")
```
## Key Findings ##

**Users are not creating manual entries**

* Only 33 distinct user id's in daily activity, 24 distinct in sleep log and 8 distinct in weight log.
* Therefore, Bellabeat should promote body positivity, so that users feel comfortable manually entering their weight.

**Not all users sleep with their fitbit**

* Only 24 users had sleep data. That's why Bellbeat should promote the battery longevity and comfortness of the device.

**Active time**

* Bellabeat should promote that people don't need high intensity workouts for burning calories.
* The app should promote all levels of excercise like yoga, marathon. Bellabeat continue to emphasize that all are welcome, and that the Leaf is so much more than tracking fitness.



